Tool-driven Revolutions in Archaeological Science

### 2018-11-16 analysis/paper/jcaa-paper.Rmd
some grammar knots cleared (I hope)

These concerns about how digital tools affect what we consider is and isn't archaeology intersect with the work of @gieryn1999cultural on boundary construction. In his cultural cartography of science, Gieryn examines boundaries between science generally and varieties of less authoritative non-science. He argues that the boundaries are drawn and redrawn contiuously in inflexible, historically changing and sometimes ambiguous ways. Gieryn claims that the cultural authority of science flows from boundary work in professional and political settings where scientists construct a public image for science as a source of credible knowledge, validate their work, and marginalize claims by competitors. Boundary work involves attributions of selected characteristics to science in order to distinguish it from non-scientific competitors. The increasing voluntary use and sharing of code by researchers, and demands for it by readers and peer reviewers is a form of active boundary work that is shifting the boundary between scientific and non- or less-scientific archaeology. We argue that the use of code as a research tool, and sharing of code as a public research product, are increasingly attributes that are being used to distinguish science from less scientific work.  The reason is because using code, and making available for others to inspect, enhancing the credibility of knowledge claims.

We do not believe that a Kuhnian paradigm-change seems to be happening in ecology or archaeology: Data-and method sharing behaviour has been around in archaeology before now and do not present incommensurate ways of thinking about the subject in the way that a classical paradigm-change should. What we can see, though, is that the need for methods to share research in a way that is reproducible for others has been acknowledged and that archaeologists have looked for tools to make this possible. Using scripting languages and data repositories to share the scripts is a solution that we have demonstrated enjoys increasing popularity in the archaeological research community. It has also expanded the 'trading zone' [@Galison_1997] where archaeologists collaborate with specialists in other fields to develop a shared language to get things done. In using scripting languages archaeologists are working at a trading zone with computer scientists and computational reserchers in other fields to coordinate how tools from one domain can be useful in another. We can also see archaeologists in similar trading zones when using GIS, network methods, agent-based models, isotope analysis, remote sensing, and DNA analysis to answer archaeological questions. Archaeologists' embrace of, or turn towards, these methods could also be argued to be tool-driven revolutions. However, what makes the use of code unique as a tool-driven revolution is that coding is transcendental, it is universal method with which we can do all, or be central to, of those previously-mentioned approaches. That is why we believe archaeologists should pay attention to scripted workflows regardless of their usual toolkit and methodological interests.

>>>>

These concerns about how digital tools affect what we consider is and isn't archaeology intersect with the work of @gieryn1999cultural on boundary construction. In his cultural cartography of science, Gieryn examines boundaries between science generally and varieties of less authoritative non-science. He argues that the boundaries are drawn and redrawn contiuously in flexible, historically changing and sometimes ambiguous ways. Gieryn claims that the cultural authority of science flows from boundary work in professional and political settings where scientists construct a public image for science as a source of credible knowledge, validate their work, and marginalize claims by competitors. Boundary work involves attributions of selected characteristics to science in order to distinguish it from non-scientific competitors. The increasing voluntary use and sharing of code by researchers, and demands for it by readers and peer reviewers is a form of active boundary work that is shifting the boundary between scientific and non- or less-scientific archaeology. The authors, as members of this group, argue that the use of code as a research tool, and the sharing of code as a public research product, are increasingly attributes that are being used to distinguish science from less scientific work. The reason for this is that using code, and making it available for others to inspect, enhances the credibility of knowledge claims.

We do not believe that a Kuhnian paradigm-change seems to be happening in ecology or archaeology: Data-and method sharing behaviour has been around in archaeology before now and does not present incommensurate ways of thinking about the subject in the way that a classical paradigm-change should. What we can see, though, is that the need for methods to share research in a way that is reproducible for others has been acknowledged and that archaeologists have looked for tools to make this possible. Using scripting languages and data repositories to share the scripts is a solution that we have demonstrated enjoys increasing popularity in the archaeological research community. It has also expanded the 'trading zone' [@Galison_1997] where archaeologists collaborate with specialists in other fields to develop a shared language to get things done. In using scripting languages archaeologists are working at a trading zone with computer scientists and computational reserchers in other fields to coordinate how tools from one domain can be useful in another. We can also see archaeologists in similar trading zones when using GIS, network methods, agent-based models, isotope analysis, remote sensing, and DNA analysis to answer archaeological questions. Archaeologists' embrace of, or turn towards, these methods could also be argued to be tool-driven revolutions. However, what makes the use of code unique as a tool-driven revolution is that coding is transcendental, it is an universal method with which we can do all, or which is central to any, of those previously-mentioned approaches. That is why we believe archaeologists should pay attention to scripted workflows regardless of their usual toolkit and methodological interests.
